---
title: "banana-quality-r"
author: "Kevin. T"
date: "2024-03-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## R Markdown
```{r}
banana_df <- read_csv("banana_quality.csv")
head(banana_df)
```

## Dimension Reduction 
Principle Component Analysis 

```{r}
library(ggcorrplot)
banana_num <- banana_df %>%
  select(-c(Quality))
corr_matrix <- cor(banana_num)
ggcorrplot(corr_matrix)
```



Variables are not closely correlated with exception.

Now to standardise data: 
```{r}
# I will store sweetness in a separate variable
sweetness = banana_df$Sweetness

banana_std <- banana_df %>%
  select(-c(Quality, Sweetness)) %>%
  scale() %>% tibble
```

```{r}
set.seed(123)
n = nrow(banana_std)
index <- sample(1:n,n*0.6)
banana_train <- banana_std %>%
  slice(index)
banana_valid <- banana_std %>%
  slice(-index)

sweetness_train = sweetness[index]
sweetness_valid = sweetness[-index]
```


Apply kNN with k = 3 to predict the sweetness of bananas in validation sample. 

```{r}
library(FNN)

knn_k3_pred = knn.reg(train = as.matrix(banana_train),
                      test = as.matrix(banana_valid),
                      y = as.matrix(sweetness_train),
                      k = 3)
```

```{r}
library(forecast)
round(accuracy(knn_k3_pred$pred,sweetness_valid),1)
```
Finding optimal K

```{r}
K = 1:50
RMSEall = numeric(length(K))
for (k in K){
  knn_kk_pred = knn.reg(train = as.matrix(banana_train),
                      test = as.matrix(banana_valid),
                      y = sweetness_train,
                      k=k)
  
  RMSEall[k] = accuracy(knn_kk_pred$pred, sweetness_valid)[2]
                      
}
plot(K,RMSEall, type = 'l')


```
```{r}
knn_k27_pred = knn.reg(train = as.matrix(banana_train),
                      test = as.matrix(banana_valid),
                      y = as.matrix(sweetness_train),
                      k = 27)
round(accuracy(knn_k27_pred$pred, sweetness_valid),1)
```

```{r}
RMSEall
```


```{r}
# banana_df %>%
#   select(-Quality) %>%
#   prcomp(scale. = TRUE) -> pcaout
# summary(pcaout)
# screeplot(pcaout, type = 'l')
# pca_scores <- pcaout$x[, 1:5]
# banana_pca <- cbind(banana_df, PCA1 = pca_scores[, 1], 
#                     PCA2 = pca_scores[, 2],
#                     PCA3)
# head(banana_pca)
```
7 principal components shown.
First principal component cumulative proportion 0.308, explaining only 30.8% of
total variance alone. 

Proportion of variance explained by first two PCs together 50.8%, but proportion
of variance by second PC alone is 20%. 

Proportion of variance explained by the first five PCs together is 88.69%. 





Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
