---
title: 'Ch 3: Time series decomposition'
author: "Kevin. T"
date: '2022-03-29'
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fpp3)
library(forecast)
library(tibble)
library(tsibble)
library(fabletools)
library(ggplot2)
```

## Transormations and adjustments
*Per capita adjustments*
```{r}
global_economy %>%
filter(Country == "Australia") %>%
autoplot(GDP)
```
\newpage 
```{r}
global_economy %>%
filter(Country == "Australia") %>%
autoplot(GDP / Population)
```
*Inflation adjustments* 
```{r}
print_retail <- aus_retail %>%
filter(Industry == "Newspaper and book retailing") %>%
group_by(Industry) %>%
index_by(Year = year(Month)) %>%
summarise(Turnover = sum(Turnover))

aus_economy <- global_economy %>%
filter(Code == "AUS")

print_retail %>%
left_join(aus_economy, by = "Year") %>%
mutate(Adjusted_turnover = Turnover / CPI*100) %>%
pivot_longer(c(Turnover, Adjusted_turnover),
names_to = "Type", values_to = "AUD") %>%
mutate(Type = factor(Type, levels=c("Turnover","Adjusted_turnover"))) %>%
ggplot(aes(x = Year, y = AUD)) +
geom_line() +
facet_grid(Type ~ ., scales = "free_y") +
labs(title = "Turnover: Australian print media industry", y = "$AU") 
```
*Mathematical Transformations*
```{r}
food <- aus_retail %>%
filter(Industry == "Food retailing") %>%
summarise(Turnover = sum(Turnover))

food %>% autoplot(sqrt(Turnover)) +
labs(y = "Square root turnover")

food %>% autoplot(log(Turnover)) +
labs(y = "Log turnover")

food %>% autoplot(-1/Turnover) +
labs(y = "Inverse turnover")
```
*Box-Cox Transformations*
```{r}
food %>%
features(Turnover, features = guerrero)

food %>% autoplot(box_cox(Turnover, 0.0524)) +
labs(y = "Box-Cox transformed turnover")
```

## Time series decomposition

```{r}
us_retail_employment <- us_employment %>%
filter(year(Month) >= 1990, Title == "Retail Trade") %>%
select(-Series_ID)
us_retail_employment

us_retail_employment %>%
autoplot(Employed) +
labs(y="Persons (thousands)", title="Total employment in US retail")

us_retail_employment %>%
model(stl = STL(Employed))

dcmp <- us_retail_employment %>%
model(stl = STL(Employed))
components(dcmp)

us_retail_employment %>%
autoplot(Employed, color='gray') +
autolayer(components(dcmp), trend, color='#D55E00') +
labs(y="Persons (thousands)", title="Total employment in US retail")

components(dcmp) %>% autoplot()
components(dcmp) %>% gg_subseries(season_year)
```
*Seasonal adjustment *
```{r}
us_retail_employment %>%
autoplot(Employed, color='gray') +
autolayer(components(dcmp), season_adjust, color='#0072B2') +
labs(y="Persons (thousands)", title="Total employment in US retail")
```

## Moving averages
```{r}
global_economy %>% filter(Country == "Australia") %>%
autoplot(Exports) +
labs(y="% of GDP", title= "Total Australian exports")
```
*Moving average trend-cycle *
```{r}
us_retail_employment_ma <- us_retail_employment %>%
mutate(
`12-MA` = slider::slide_dbl(Employed, mean,
.before = 5, .after = 6, .complete = TRUE),
`2x12-MA` = slider::slide_dbl(`12-MA`, mean,
.before = 1, .after = 0, .complete = TRUE)
)
us_retail_employment_ma %>%
autoplot(Employed, color = "gray") +
autolayer(us_retail_employment_ma, vars(`2x12-MA`),
color = "#D55E00") +
labs(y = "Persons (thousands)",
title = "Total employment in US retail")
```
##Classical decomposition
```{r}
us_retail_employment %>%
model(classical_decomposition(Employed, type = "additive")) %>%
components() %>%
autoplot() + xlab("Year") +
ggtitle("Classical additive decomposition of total
US retail employment")

```
*STL decomposition*

```{r}
us_retail_employment %>%
model(STL(Employed ~ season(window=9), robust=TRUE)) %>%
components() %>% autoplot() +
labs(title = "STL decomposition: US retail employment")

```

```{r}
us_retail_employment %>%
model(STL(Employed ~ season(window=5))) %>%
components()
us_retail_employment %>%
model(STL(Employed ~ trend(window=15) +
season(window="periodic"),
robust = TRUE)
) %>% components()
```

```{r}
us_retail_employment %>% model(STL(Employed)) %>%
components() %>% autoplot()
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
