---
title: "Forecasting turnover of Victoria's cafe and fast-food services"
author: "Kevin. T"
date: "2024-01-26"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(forecast)
library(zoo)
library(fpp3)
library(readr)
```

## Introduction

Data source: https://explore.data.abs.gov.au/vis?tm=turnover&pg=0&df[ds]=INDUSTRY_TOPICS&df[id]=RT&df[ag]=ABS&df[vs]=1.0.0&pd=2022-07%2C&dq=.20%2B41%2B42%2B43%2B44%2B45%2B46..AUS.M&ly[cl]=INDUSTRY&ly[rw]=TIME_PERIOD&ly[rs]=MEASURE%2CTSEST 



```{r}
fs_turnover <- read_csv("vic-food-turnover.csv")
head(fs_turnover)

```

### Data Cleaning 
```{r}
 fs_turnover<- fs_turnover %>%
  rename(Turnover = OBS_VALUE_VIC, Month = TIME_PERIOD) %>%
  select(Month, Turnover) %>%
  mutate(Month = yearmonth(Month)) %>%
  as_tsibble(index = Month)

head(fs_turnover)

```
```{r}
fs_turnover %>% autoplot(Turnover) + ggtitle("Turnovers of Victorian cafes, restaurants & fast-food") +
  ylab("$AUD (millions)") 
```

### Exploring time series 
```{r}
# seasonal plot 
fs_turnover %>% 
  filter(year(Month)>2018) %>%
  gg_season(Turnover, labels = 'both') + guides(colour = "none") +
  labs(
    title = "Seasonal Plot ",
    y = "$AUD (millions)")

#subseries plot
fs_turnover %>%
  filter(year(Month)>2018) %>%
  gg_subseries(Turnover) +
  labs(
    title = "Turnover: Western Australia Department Stores Seasonal Plot 1982-2022 ",
    y = "$AUD (millions)"
  )
```

## Exploring Forecasting Models

```{r}
#training 1982 - 2019
fs_train <- fs_logtransformed %>% slice(1:453)
#test 2020- present 
fs_test <- fs_logtransformed %>%
  filter(year(Month) >= 2020)
```


```{r}
fs_turnover %>% autoplot(Turnover) + ggtitle("Turnovers of Victorian cafes, restaurants & fast-food") +
  ylab("$AUD (millions)") 

fs_turnover %>%
  features(Turnover, features = guerrero) 

fs_logtransformed <- fs_turnover %>%
  mutate(log_turnover = log(Turnover))

fs_logtransformed %>% autoplot(log_turnover) + ggtitle("LOG Turnovers of Victorian cafes, restaurants & fast-food") +
  ylab("$AUD (millions)") 
```


### Evaluating Seasonal Naive vs drift model  
```{r}
fit_train <- fs_train %>%
  model(
    snaive = SNAIVE(log(Turnover)),
    drift = RW(log(Turnover) ~ drift())
  )
fit_train %>%  glance()

fc_test <- fit_train %>%
  forecast(h = "2 years")
fc_test %>% autoplot(fs_turnover, alpha=0.6)

fit_train %>% accuracy() %>% select(MAE, RMSE, MAPE, MASE)
fc_test %>% accuracy(fs_turnover) %>% select(MAE, RMSE, MAPE, MASE)
```
```{r}
fit_snaive <- fs_turnover %>%
  model(
    snaive = SNAIVE(log(Turnover)),
  )
fit_snaive %>%
  gg_tsresiduals()
```
### Generating Forecast with SNAIVE model 

```{r}
turnover_benchmark_fit <-  fs_train %>%
  model(Seasonal_naive = SNAIVE(Turnover))

#forecast 3 years from dec 2019 since our data ends on november 2023. 
turnover_benchmark_fc <- turnover_benchmark_fit %>%
  forecast(h = '4 years')  

turnover_benchmark_fc %>% autoplot(fs_turnover) +
  labs(title = "SNAIVE: Turnover for Food Services in Victoria",
       y = "$AUD (millions)") +
  guides(colour = guide_legend(title = "Forecast"))


turnover_benchmark_fc %>% autoplot(fs_test) +
  labs(title = "SNAIVE: Turnover forecast for Food Services in Victoria",
       y = "$AUD (millions)") +
  guides(colour = guide_legend(title = "Forecast"))

benchmark_point_intervals <- fs_train%>%
  model(SNAIVE(Turnover)) %>%
  forecast(h = '4 years') %>%
  hilo()
print(benchmark_point_intervals)
```
### Evaluatig Exponential Smoothing (ETS) 


```{r}
fit_ets <- fs_turnover %>%
  model(ETS(log(Turnover)))
report(fit_ets)

```
Shows no trend, but our plot clearly shows a steady upward trend (with the exception of the COVID pandemic)

```{r}
components(fit_ets) %>% autoplot()
```

```{r}
#ETS-  MAM c
ETS_fit <- fs_train %>%
  model(
    hw_multi = ETS(log(Turnover) ~ error("M") + trend("A") + season("M")),
    no_trend = ETS(log(Turnover) ~ error("M") + trend("N") + season("M")))
report(ETS_fit)
```

```{r}
ETS_fit <- fs_train %>%
  model(
    ETS(log(Turnover) ~ error("M") + trend("A") + season("M")))
ETS_fc <- ETS_fit %>%
  forecast(h = '4 years')

ETS_fc  %>% autoplot(filter(fs_turnover, year(Month)>2019))  +  
  labs(title = "ETS(M,A,M) : Turnover for Food Services in Victoria",
       y = "$AUD (millions)") +
  guides(colour = guide_legend(title = "Forecast"))

ETS_point_intervals <- ETS_fit %>%
  forecast(h= '4 years') %>%
  hilo()
print(ETS_point_intervals)
```

### ARIMA

Further transformation  
```{r}
fs_turnover %>% autoplot(log(Turnover)  %>% 
  difference(12))

#Automatic difference selection
fs_turnover %>% mutate(log_turnover = log(Turnover)) %>%
  features(log_turnover, unitroot_nsdiffs) 

fs_turnover %>% autoplot(log(Turnover)  %>% 
  difference(12) %>% difference (1)) 



```


```{r}

fs_transformed %>% gg_tsdisplay(log(Turnover) %>% difference(12) %>% difference(), 
                           plot_type = 'partial')


```

Observed:
* non-seasonal spikes in the PACF: AR(2)
* seasonal MA(1) due to the large spike at 12 in the ACF
* Model Chosen: 

```{r}
arima_210_011 <- fs_turnover %>% model(
  ARIMA(log(Turnover)~pdq(2,1,0)+PDQ(0,1,1))
  )
arima_210_011 %>% report()

arima_210_011 %>% gg_tsresiduals()

augment(arima_210_011) %>% features(.innov, ljung_box, lag=24, dof=3)

augment(arima_210_011) %>% gg_tsdisplay(.innov, plot_type = 'partial')

```




Other models worth considering:
* MA(2): 2 spikes in ACF plot: ARIMA(0,1,2)(0,1,1)
* AR(4): large spike in ACF: ARIMA(4,1,0)(0,1,1)
* AR(2)MA(2): ARIMA(2,1,2)(0,1,1)


```{r}
arima.models <- fs_turnover %>% model(
  arima_210_011 = ARIMA(log(Turnover)~pdq(2,1,0)+PDQ(0,1,1)),
  arima_012_011 = ARIMA(log(Turnover)~pdq(0,1,2)+PDQ(0,1,1)),
  arima_410_011 = ARIMA(log(Turnover)~pdq(4,1,0)+PDQ(0,1,1)),
  arima_212_011 = ARIMA(log(Turnover)~pdq(2,1,2)+PDQ(0,1,1))
  )
arima.models %>%  glance() %>% select(.model, AIC, AICc, BIC)


```

Lower values of AIC, AICc, and BIC are preferred.
Comparing the models based on these criteria, ARIMA(0,1,2) (arima_012_011) has the lowest AIC, AICc, and BIC values among the listed models.


The auto.arima() function is useful, but anything automated can be a little dangerous, and it is worth understanding

```{r}

arima_models_auto <- fs_turnover %>%
  model(arima_auto1 = ARIMA(log(Turnover)),
        arima_auto2 = ARIMA(log(Turnover), stepwise = FALSE, approximation = FALSE)
  )
arima_models_auto %>% select(arima_auto1) %>%  report()


```

The auto funcitonchose an ARIMA(0,1,4)(2,0,0)[12] for us. 

```{r}
arima_models_auto %>% select(arima_auto2) %>%  report()

```
### Choosing overlal model;

```{r}
fit_all <- fs_turnover %>%
  model(
    arima_210_011 = ARIMA(log(Turnover)~pdq(2,1,0)+PDQ(0,1,1)),
  arima_012_011 = ARIMA(log(Turnover)~pdq(0,1,2)+PDQ(0,1,1)),
  arima_410_011 = ARIMA(log(Turnover)~pdq(4,1,0)+PDQ(0,1,1)),
  arima_212_011 = ARIMA(log(Turnover)~pdq(2,1,2)+PDQ(0,1,1)),
  arima_212_200 = ARIMA(log(Turnover), stepwise = FALSE, approx = FALSE)
  )
glance(fit_all)

```
ARIMA(2,1,2)(2,0,0)

```{r}
arima_models_auto %>%
  select(arima_auto2) %>%
  forecast() %>%
  autoplot(fs_turnover)

```

```{r}
arima_models_auto %>%
  select(arima_auto2) %>%
  forecast() %>%
  autoplot(filter(fs_turnover,year(Month)>=2020))

```
### Other Forecasting 


```{r}
#training 1982 - 2019
fs_train <- fs_turnover %>% slice(1:453)
#test 2020-2022
fs_test <- fs_turnover %>%
  filter(year(Month) >= 2020)
```


```{r}

#ARIMA - best arima model ARIMA(2,1,2)(0,1,1)
fs_train %>% autoplot(log(Turnover)  %>% 
  difference(12) %>% difference (1))


arima_fit <- fs_train %>%
  model(
    arima012111 = ARIMA(log(Turnover) ~ pdq(2,1,2) + PDQ(0,1,1))
  )

arima_fc <- arima_fit %>%
  select(arima012111) %>%
  forecast(h =47)

arima_fc %>% autoplot(filter(fs_turnover, year(Month)>2019)) +
   labs(title = "ARIMA(2,1,2)(0,1,1): Turnover for Department Stores in Western Australia",
       y = "$AUD (millions)") +
  guides(colour = guide_legend(title = "Forecast"))

arima_point_intervals <- arima_fit %>%
  forecast(h= '3 years') %>%
  hilo() 
arima_point_intervals
print(arima_point_intervals)


```


```{r}
models_final <- fs_train %>% model(
  snaive = SNAIVE(log(Turnover)),
  ETS_MAdM=ETS(Turnover~error("M")+trend("Ad")+season("M")),
  ARIMA_014_011=ARIMA(log(Turnover)~pdq(0,1,4)+PDQ(0,1,1))
)
models_final %>%
  forecast(h=48) %>%
  autoplot(fs_turnover %>% filter(year(Month)>=2020), alpha=0.8)
```


```{r}
accuracy(turnover_benchmark_fc,fs_test) %>% select(.model,.type, ME, RMSE,MAPE, MASE, RMSSE)
accuracy(ETS_fc, fs_test) %>% select(.model,.type, ME, RMSE,MAPE, MASE, RMSSE)
accuracy(arima_fc, fs_test) %>% select(.model,.type, ME, RMSE,MAPE, MASE, RMSSE)
```

```{r}

```

```{r}
models_final %>%
  forecast(h='3 years') %>%
  accuracy(fs_test) %>% select(.model,.type, ME, RMSE,MAPE, MASE, RMSSE)

```

```{r}
all_model_fit <- fs_turnover %>%
  model(
    Seasonal_naive = SNAIVE(Turnover),
    arima012111 = ARIMA(log(Turnover) ~ pdq(2,1,2) + PDQ(1,1,1)),
    hw_multi = ETS(Turnover ~ error("M") + trend("A") + season("M"))
        )
all_model_fc <- all_model_fit %>%
  forecast(h=24)

all_model_fc  %>% 
  autoplot(filter(fs_turnover, year(Month)>=2020), alpha=0.6)
```

